---
title: "R Notebook"
output: html_notebook
---
#Libraries
```{r}
library("phyloseq")
library("ggplot2")
library(DESeq2)
```

#Importing files for microbiome analysis with phyloseq
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in otu table
otu_table = read.csv("otu_table_97.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table,rownames=TRUE)
#read in taxonomy
taxonomy = read.csv("taxonomy_97.csv",sep=",",row.names=1)
taxonomy=as.matrix(taxonomy)
#read in metadata
metadata=read.table("metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree_97.nwk")
```

##Generating a phyloseq object 
```{r}
#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
META$Phase=as.factor(META$Phase)
META$STEC=as.factor(META$STEC)
META$STEC=as.factor(META$STEC)
META$STEC_prev_group <- factor(META$STEC_prev_group, levels = c("Low","High"))

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)
taxa_names(phy_tree)

# Same sample names
sample_names(OTU)
sample_names(META)

#(tree was already imported as a phyloseq object)
physeq=phyloseq(OTU,TAX,META,phy_tree)
```

#Filtering samples and taxa for further analyses
```{r}
physeq_two = subset_taxa(physeq, Order != "Chloroplast")
physeq_three = subset_taxa(physeq_two, Family != "Mitochondria")
physeq_hindgut <- subset_samples(physeq_three, Type%in%c("RAJ","FG"))
physeq_STEC <- subset_samples(physeq_hindgut, STEC%in%c("1","0"))
physeq_dairy <- subset_samples(physeq_STEC, Herdtype%in%c("Dairy"))

physeq = physeq_dairy
```

#OTU
#PHYLUM
```{r}
library(tidyverse)

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/OTUabundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/metadata_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/metadata_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant Phylum STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_phylum.csv") #Writing `sigtab` to
```

##Volcano plot Phylum STEC dairy
```{r}
 # if (!requireNamespace('BiocManager', quietly = TRUE))
  #  install.packages('BiocManager')
   # BiocManager::install('EnhancedVolcano')

library(EnhancedVolcano)

phylum_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.01,
    FCcutoff = 0.2,
    y = 'pvalue',
    labvjust = -1,
    xlim = c(-1,1),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Phylum")

phylum_plot
```

```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 7

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_phylum.pdf',width=setWidth,height=6,pointsize=setFontSize)
phylum_plot
dev.off()

```
#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant Family STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_family.csv") #Writing `sigtab` to
```

##Volcano plot Family STEC dairy
```{r fig.width=10, fig.height=10}

family_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Family",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

family_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 8

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_family.pdf',width=setWidth,height=10,pointsize=setFontSize)
family_plot
dev.off()

```

#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant Genus STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_genus.csv") #Writing `sigtab` to
```

##Volcano plot Genus STEC dairy
```{r fig.width=10, fig.height=10}

genus_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Genus",
    labSize = 4,
   labFace = "italic",
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

genus_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_genus.pdf',width=setWidth,height=10,pointsize=setFontSize)
genus_plot
dev.off()

```

#CLASS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_class <- tax_glom(physeq, taxrank = 'Class')

#Transform abundance table to data-frame
df_class <- psmelt(glom_class)

#Group ASVs at the genus level
grouped_class <- df_class %>%
  select(Sample, Class, Abundance)

class_matrix = reshape(grouped_class,idvar="Sample",timevar = "Class", times = "Abundance",direction="wide")


write.table(class_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/class_abundance_dairy_STEC.csv", sep = ',', col.names = TRUE, row.names = F)
```

##CLASS objects
```{r}
#abundance table by sample
cts_class=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/class_abundance_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_class <- t(cts_class)
cts_class <-  as.data.frame(cts_class)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis class STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_class,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant class STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_class.csv") #Writing `sigtab` to
```

##Volcano plot class STEC dairy
```{r fig.width=10, fig.height=5}

class_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Genus",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

class_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_class.pdf',width=setWidth,height=5,pointsize=setFontSize)
class_plot
dev.off()

```

#ORDER
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_order <- tax_glom(physeq, taxrank = 'Order')

#Transform abundance table to data-frame
df_order <- psmelt(glom_order)

#Group ASVs at the genus level
grouped_order <- df_order %>%
  select(Sample, Order, Abundance)

order_matrix = reshape(grouped_order,idvar="Sample",timevar = "Order", times = "Abundance",direction="wide")


write.table(order_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/order_abundance_dairy_STEC.csv", sep = ',', col.names = TRUE, row.names = F)
```

##CLASS objects
```{r}
#abundance table by sample
cts_order=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/order_abundance_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_order <- t(cts_order)
cts_order <-  as.data.frame(cts_order)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis class STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_order,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant order STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_order.csv") #Writing `sigtab` to
```

##Volcano plot class STEC dairy
```{r fig.width=10, fig.height=5}

order_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Order",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

order_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_class.pdf',width=setWidth,height=5,pointsize=setFontSize)
class_plot
dev.off()

```

#ANALYSIS WITH FARM AS RANDOM EFFECT
#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/metadata_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_dairy.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~STEC_pos + Farm + STEC_pos:Farm)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_posSTEC.Farm4D")
```

##Export table with differentially abundant Phylum STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_phylum.csv") #Writing `sigtab` to
```

##Volcano plot Phylum STEC dairy
```{r}
 # if (!requireNamespace('BiocManager', quietly = TRUE))
  #  install.packages('BiocManager')
   # BiocManager::install('EnhancedVolcano')

library(EnhancedVolcano)

phylum_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.01,
    FCcutoff = 0.2,
    y = 'pvalue',
    labvjust = -1,
    xlim = c(-1,1),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Phylum")

phylum_plot
```

```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 7

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_phylum.pdf',width=setWidth,height=6,pointsize=setFontSize)
phylum_plot
dev.off()

```
#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant Family STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_family.csv") #Writing `sigtab` to
```

##Volcano plot Family STEC dairy
```{r fig.width=10, fig.height=10}

family_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Family",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

family_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 8

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_family.pdf',width=setWidth,height=10,pointsize=setFontSize)
family_plot
dev.off()

```

#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant Genus STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_genus.csv") #Writing `sigtab` to
```

##Volcano plot Genus STEC dairy
```{r fig.width=10, fig.height=10}

genus_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Genus",
    labSize = 4,
   labFace = "italic",
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

genus_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_genus.pdf',width=setWidth,height=10,pointsize=setFontSize)
genus_plot
dev.off()

```

#CLASS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_class <- tax_glom(physeq, taxrank = 'Class')

#Transform abundance table to data-frame
df_class <- psmelt(glom_class)

#Group ASVs at the genus level
grouped_class <- df_class %>%
  select(Sample, Class, Abundance)

class_matrix = reshape(grouped_class,idvar="Sample",timevar = "Class", times = "Abundance",direction="wide")


write.table(class_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/class_abundance_dairy_STEC.csv", sep = ',', col.names = TRUE, row.names = F)
```

##CLASS objects
```{r}
#abundance table by sample
cts_class=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/class_abundance_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_class <- t(cts_class)
cts_class <-  as.data.frame(cts_class)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis class STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_class,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant class STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_class.csv") #Writing `sigtab` to
```

##Volcano plot class STEC dairy
```{r fig.width=10, fig.height=5}

class_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Genus",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

class_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_class.pdf',width=setWidth,height=5,pointsize=setFontSize)
class_plot
dev.off()

```

#ORDER
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_order <- tax_glom(physeq, taxrank = 'Order')

#Transform abundance table to data-frame
df_order <- psmelt(glom_order)

#Group ASVs at the genus level
grouped_order <- df_order %>%
  select(Sample, Order, Abundance)

order_matrix = reshape(grouped_order,idvar="Sample",timevar = "Order", times = "Abundance",direction="wide")


write.table(order_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/order_abundance_dairy_STEC.csv", sep = ',', col.names = TRUE, row.names = F)
```

##CLASS objects
```{r}
#abundance table by sample
cts_order=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/order_abundance_dairy_STEC.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_order <- t(cts_order)
cts_order <-  as.data.frame(cts_order)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis class STEC dairy
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_order,
                              colData = coldata,
                              design= ~STEC_pos)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="STEC_pos_STEC_vs_Control")
```

##Export table with differentially abundant order STEC dairy
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_dairy_STEC_order.csv") #Writing `sigtab` to
```

##Volcano plot class STEC dairy
```{r fig.width=10, fig.height=5}

order_plot <-  EnhancedVolcano(res,
    lab = rownames(res),
    x = 'log2FoldChange',
    pCutoff = 0.05,
    FCcutoff = 0.5,
    y = 'pvalue',
   labvjust = 1.2,
    labhjust = 2,
    xlim = c(-5,5),
    ylim = c(0,5),
    title = "STEC shedders vs. non shedders",
    subtitle = "Order",
    labSize = 4,
    drawConnectors = T,
    widthConnectors = .1,
   typeConnectors = open,
   arrowheads = F)

order_plot
```
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='volcano_dairy_STEC_class.pdf',width=setWidth,height=5,pointsize=setFontSize)
class_plot
dev.off()

```
