---
title: "R Notebook"
output: html_notebook
---
---
title: "R Notebook"
output: html_notebook
---
#Libraries
```{r}
library("phyloseq")
library("ggplot2")
library(DESeq2)
```

#Importing files for microbiome analysis with phyloseq
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in otu table
otu_table = read.csv("otu_table_97.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table,rownames=TRUE)
#read in taxonomy
taxonomy = read.csv("taxonomy_97.csv",sep=",",row.names=1)
taxonomy=as.matrix(taxonomy)
#read in metadata
metadata=read.table("metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree_97.nwk")
```

##Generating a phyloseq object 
```{r}
#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
META$Phase=as.factor(META$Phase)
META$STEC=as.factor(META$STEC)
META$STEC=as.factor(META$STEC)
META$STEC_prev_group <- factor(META$STEC_prev_group, levels = c("Low","High"))

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)
taxa_names(phy_tree)

# Same sample names
sample_names(OTU)
sample_names(META)

#(tree was already imported as a phyloseq object)
physeq=phyloseq(OTU,TAX,META,phy_tree)
```

#Filtering samples and taxa for further analyses
```{r}
physeq_two = subset_taxa(physeq, Order != "Chloroplast")
physeq_three = subset_taxa(physeq_two, Family != "Mitochondria")
physeq_hindgut <- subset_samples(physeq_three, Type%in%c("RAJ","FG"))
physeq_forage <- subset_samples(physeq_hindgut,!is.na(Pforage))
physeq_TEMP <- subset_samples(physeq_hindgut,!is.na(TempC_avg_5D))
physeq_STEC <- subset_samples(physeq_hindgut, STEC%in%c("1","0"))
physeq_DIM <- subset_samples(physeq_hindgut,!is.na(DIM))
physeq_LAC <- subset_samples(physeq_hindgut,!is.na(LACT))

```

#FORAGE
```{r}
physeq = physeq_forage
```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_all.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects/metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

coldata <- subset(coldata, Type%in%c("RAJ","FG"))
coldata <- subset(coldata, !is.na(Pforage))

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_all.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum PFORAGE
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~Pforage)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Pforage")
```

##Export table with differentially abundant Phylum PFORAGE
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.07
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_pforage_phylum.csv") #Writing `sigtab` to
```

#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_forage.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_forage.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family Pforage
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~Pforage)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Pforage")
```

##Export table with differentially abundant Family forage
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_pforage_family.csv") #Writing `sigtab` to
```


#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_pforage.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_pforage.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus Pforage
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~Pforage)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Pforage")
```

##Export table with differentially abundant Genus Pforage
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_pforage_genus.csv") #Writing `sigtab` to
```

#TEMPERATURE
```{r}
physeq = physeq_TEMP 
```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_TEMP.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects/metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

coldata <- subset(coldata, Type%in%c("RAJ","FG"))
coldata <- subset(coldata, !is.na(TempC_avg_5D))

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_TEMP.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum TEMP
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~TempC_avg_5D)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="TempC_avg_5D")
```

##Export table with differentially abundant TEMP
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.07
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_TEMP_phylum.csv") #Writing `sigtab` to
```

#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_TEMP.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_TEMP.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family TEMP
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~TempC_avg_5D)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="TempC_avg_5D")
```

##Export table with differentially abundant Family TEMP
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_TEMP_family.csv") #Writing `sigtab` to
```


#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_TEMP.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_TEMP.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus TEMP
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~TempC_avg_5D)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="TempC_avg_5D")
```

##Export table with differentially abundant Genus TEMP
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_TEMP_genus.csv") #Writing `sigtab` to
```

#LACT
```{r}
physeq = physeq_LAC 
```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_LACT.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects/metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

coldata <- subset(coldata, Type%in%c("RAJ","FG"))
coldata <- subset(coldata, !is.na(LACT))

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_LACT.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum LACT
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~LACT)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="LACT")
```

##Export table with differentially abundant Phylum LACT
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.07
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_LACT_phylum.csv") #Writing `sigtab` to
```

#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_LACT.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_LACT.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family LACT
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~LACT)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="LACT")
```

##Export table with differentially abundant Family LACT
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_LACT_family.csv") #Writing `sigtab` to
```


#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_LACT.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_LACT.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus LACT
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~LACT)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="LACT")
```

##Export table with differentially abundant Genus LACT
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_LACT_genus.csv") #Writing `sigtab` to
```

#DIM
```{r}
physeq = physeq_DIM
```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_DIM.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects/metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

coldata <- subset(coldata, Type%in%c("RAJ","FG"))
coldata <- subset(coldata, !is.na(DIM))

#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_DIM.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~DIM)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="DIM")
```

##Export table with differentially abundant Phylum DIM
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.07
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_DIM_phylum.csv") #Writing `sigtab` to
```

#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_DIM.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_DIM.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~DIM)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="DIM")
```

##Export table with differentially abundant Family DIM
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_DIM_family.csv") #Writing `sigtab` to
```


#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_DIM.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_DIM.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~DIM)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="DIM")
```

##Export table with differentially abundant Genus DIM
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_DIM_genus.csv") #Writing `sigtab` to
```

#Clusters 8B
```{r}
physeq_8B <- subset_samples(physeq_hindgut,Farm%in%c("8B"))
physeq <- physeq_8B

```

#PHYLUM
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_phy <- tax_glom(physeq, taxrank = 'Phylum')

#Transform abundance table to data-frame
df_phy <- psmelt(glom_phy)

#Group ASVs at the genus level
grouped_phylum <- df_phy %>%
  select(Sample, Phylum, Abundance)

phylum_matrix = reshape(grouped_phylum,idvar="Sample",timevar = "Phylum", times = "Abundance",direction="wide")


write.table(phylum_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_8B.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Phylum objects
```{r}
#metadata
coldata=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/metadata_8B.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 


#abundance table by sample
cts_phylum=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/phylum_abundance_8B.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_phylum <- t(cts_phylum)
cts_phylum <-  as.data.frame(cts_phylum)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Phylum DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_phylum,
                              colData = coldata,
                              design= ~Phase_cluster)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Phase_cluster_Cluster2_vs_Cluster1")
```

##Export table with differentially abundant Phylum DIM
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.07
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_8B_phylum.csv") #Writing `sigtab` to
```

#FAMILY
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_fam <- tax_glom(physeq, taxrank = 'Family')

#Transform abundance table to data-frame
df_fam <- psmelt(glom_fam)

#Group ASVs at the genus level
grouped_family <- df_fam %>%
  select(Sample, Family, Abundance)

family_matrix = reshape(grouped_family,idvar="Sample",timevar = "Family", times = "Abundance",direction="wide")


write.table(family_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_8B.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Family objects
```{r}
#abundance table by sample
cts_family=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/family_abundance_8B.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_family <- t(cts_family)
cts_family <-  as.data.frame(cts_family)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Family DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_family,
                              colData = coldata,
                              design= ~Phase_cluster)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Phase_cluster_Cluster2_vs_Cluster1")
```

##Export table with differentially abundant Family DIM
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_8B_family.csv") #Writing `sigtab` to
```


#GENUS
```{r}
library(tidyverse)

#Group ASV abundances at genus level
glom_gen <- tax_glom(physeq, taxrank = 'Genus')

#Transform abundance table to data-frame
df_gen <- psmelt(glom_gen)

#Group ASVs at the genus level
grouped_genus <- df_gen %>%
  select(Sample, Genus, Abundance)

genus_matrix = reshape(grouped_genus,idvar="Sample",timevar = "Genus", times = "Abundance",direction="wide")


write.table(genus_matrix, file = "/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_8B.csv", sep = ',', col.names = TRUE, row.names = F)
```

##Genus objects
```{r}
#abundance table by sample
cts_genus=read.table("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/genus_abundance_8B.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 

cts_genus <- t(cts_genus)
cts_genus <-  as.data.frame(cts_genus)
#coldata <- as.data.frame(coldata)
```

##DESEQ2 analysis Genus DIM
```{r}
#design = model, control by ~batch and measure effect by condition. batch and condition are columns in coldata
dds <- DESeqDataSetFromMatrix(countData = cts_genus,
                              colData = coldata,
                              design= ~Phase_cluster)
dds <- DESeq(dds)
resultsNames(dds) # lists the coefficients

res <- results(dds, name="Phase_cluster_Cluster2_vs_Cluster1")
```

##Export table with differentially abundant Genus 8B
```{r}
#res = results(physeq_deseq_test, cooksCutoff = FALSE)
alpha = 0.1
sigtab = res[which(res$padj<alpha), ]

write.csv(sigtab, file="/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021/deseq2_8B_genus.csv") #Writing `sigtab` to
```
