---
title: "R Notebook"
output: html_notebook
---
#Libraries
```{r}
library("phyloseq")
library("ggplot2")
library("ape")
library("gridExtra")
library("ggpubr")
library("ggsignif")
library(forcats)
library(vegan)
library(viridis)
library(cowplot)
```

#Importing files for microbiome analysis with phyloseq
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in otu table
otu_table = read.csv("otu_table_97.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table,rownames=TRUE)
#read in taxonomy
taxonomy = read.csv("taxonomy_97.csv",sep=",",row.names=1)
taxonomy=as.matrix(taxonomy)
#read in metadata
metadata=read.table("metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree_97.nwk")
```

##Generating a phyloseq object 
```{r}
#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
META$Phase=as.factor(META$Phase)
META$STEC=as.factor(META$STEC)
META$STEC=as.factor(META$STEC)
META$STEC_prev_group <- factor(META$STEC_prev_group, levels = c("Low","High"))

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)
taxa_names(phy_tree)

# Same sample names
sample_names(OTU)
sample_names(META)

#(tree was already imported as a phyloseq object)
physeq=phyloseq(OTU,TAX,META,phy_tree)
```

#Filtering samples and taxa for further analyses
```{r}
physeq_two = subset_taxa(physeq, Order != "Chloroplast")
physeq_three = subset_taxa(physeq_two, Family != "Mitochondria")
physeq_hindgut <- subset_samples(physeq_three, Type%in%c("RAJ","FG"))
```

#HISTOGRAMS TAXA ABUNDANCE

#Converting physeq to data-frame
```{r}
glom_fam <- tax_glom(physeq_hindgut, taxrank = 'Family')
glom_fam_tra  = transform_sample_counts(glom_fam, function(x) x * 100 / sum(x) )
glom_fam # should list # taxa as # phyla
data_fam <- psmelt(glom_fam) # create dataframe from phyloseq object
```

#Converting family minorities to a single category (only keep the first 11 most abundant Families)
```{r}
data_fam$Family <- as.character(data_fam$Family)
data_fam$Family[data_fam$Family != c("Ruminococcaceae", "Rikenellaceae","Prevotellaceae","Lachnospiraceae", "Bacteroidaceae", "Muribaculaceae","Christensenellaceae","Spirochaetaceae","Bacteroidales RF16 group","Bacteroidales.uncultured","Mollicutes RF39.uncultured")] <- "Minorities"
data_fam$Family <- as.factor(data_fam$Family)
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")
write.csv(data_fam,file="family_abundance_r.csv")
```

#1. Import matrix table with proportions, convert to data frame and plot

##1.1. Phylum hindgut by breed
```{r}
library(reshape)

#### PHYLUM GUT BREED###
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots/Taxa_abundance")
Phylum_breed <- read.csv("Phylum_Breed.csv", header=T)
Phylum_breed2 =melt(Phylum_breed,id=colnames(Phylum_breed)[1])


# Stacked Percent
pgb = Phylum_breed2 %>% 
  mutate(Phylum = fct_relevel(Phylum           ,"Firmicutes","Bacteroidetes","Proteobacteria","Spirochaetes","Tenericutes","Minorities")) %>%
  ggplot(aes(x=variable, y=value,fill=Phylum)) + 
  geom_bar(stat="identity", position="fill") + 
  scale_color_brewer(palette="Set3") + 
  labs(x = "Breed", y = "Relative abundance") + 
  theme_bw() +  scale_fill_brewer(palette="Paired") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

##1.2. Phylum hindgut by farm
```{r}
library(reshape)

#### PHYLUM GUT FARM###
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots/Taxa_abundance")
Phylum_farm <- read.csv("Phylum_farms.csv", header=T, check.names = F)
Phylum_farm2 =melt(Phylum_farm,id=colnames(Phylum_farm)[1])

# Stacked Percent
pgf = Phylum_farm2 %>% 
  mutate(Phylum = fct_relevel(Phylum           ,"Firmicutes","Bacteroidetes","Proteobacteria","Spirochaetes","Tenericutes","Minorities")) %>%
  ggplot(aes(x=variable, y=value,fill=Phylum)) +
  geom_bar(stat="identity", position="fill") +
  labs(x = "Farm", y = "Relative abundance") + 
  theme_bw() +  
  scale_fill_brewer(palette="Set2") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_text(colour = "black")) +
  theme(legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5))
pgf
```

##1.2. Family hindgut by farm
```{r}
library(reshape)
#### PHYLUM GUT FARM###
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots/Taxa_abundance")
taxa <- read.csv("Family_hindgut_farms.csv", header=T,check.names=F)
taxa_df =melt(taxa,id=colnames(taxa)[1])
# Stacked Percent
hist_fam = taxa_df %>% 
  mutate(Family = fct_relevel(Family,"Ruminococcaceae", "Rikenellaceae","Prevotellaceae","Lachnospiraceae", "Bacteroidaceae", "Muribaculaceae","Christensenellaceae","Spirochaetaceae","Bacteroidales RF16 group","Bacteroidales uncultured","Mollicutes RF39 uncultured", "Minorities")) %>%
  ggplot(aes(x=variable, y=value,fill=Family)) + 
  geom_bar(stat="identity", position="fill") + 
  scale_color_brewer(palette="Set3") + 
  theme(plot.subtitle = element_text(vjust = 1), plot.caption = element_text(vjust = 1)) +
  labs(x = "Farm", y = "Relative abundance") + 
  theme_bw() +  
  scale_fill_brewer(palette="Paired") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.text.x = element_text(colour = "black")) +
  theme(legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-5,-5,-5,-5))
hist_fam
```

#Phylum breed and farm gut
```{r}
Taxa_gut = ggarrange(pgf,hist_fam, nrow = 1, ncol = 2, align="h",labels = c("A","B"))
Taxa_gut
```

#3. Histogram with x-axis corresponding to samples (instead of farms)

#Importing metadata to merge with a customized table 
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in metadata
meta=read.table("metadata_up.csv", sep = ",", header = TRUE, na.strings = "") 
meta
```

#Importing matrix with the values, merging with metadata and customizing order for the plot
```{r}
library(reshape)
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots/Taxa_abundance")

rel_abd_mx <- read.csv("Sample_family.csv", header=T)

rel_abd_df =melt(rel_abd_mx,id=colnames(rel_abd_mx)[1]) #Converting matrix to df

rel_abd_df = dplyr::rename(rel_abd_df, Family = variable) #Renaming variable by Family

taxa_meta <- merge(rel_abd_df,meta,by=c("ID")) #Merging df with medata

taxa_meta$Farm_STEC = NA #Create new column

taxa_meta$Farm = factor(taxa_meta$Farm, levels=c('1B','4D','2D','9D','8B')) #New order for the facet histogram

#taxa_meta$Farm_STEC = paste(taxa_meta$Farm,taxa_meta$STEC_pos) #New column with concatenated values

#taxa_meta$Farm_STEC = factor(taxa_meta$Farm_STEC, levels=c('1B STEC','1B Control', '1B ND','4D STEC','4D Control', '4D ND','2D STEC','2D Control', '2D ND','9D STEC','9D Control', '9D ND','8B STEC','8B Control', '8B ND')) #New order for the facet histogram. I did not use it because there are too many categories.
```

#I created a new table in Excel with the customized sample order. I ordered them by farm and then by proportion of Ruminococcaceae in descending order.
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots/Taxa_abundance")

sample_order <- read.csv("sample_order_des.csv", header=T)

#sample_order <- read.csv("sample_ru_order.csv", header=T)

as.data.frame(sample_order)

taxa_meta$ID = factor(taxa_meta$ID, levels = sample_order$ID) #Customized order of the samples
```
#SAMPLE HISTOGRAM
```{r}
# New facet label names for supp variable
supp.labs <- c("Crossbreed 1B", "Holstein 4D", "Holstein 2D", "Holstein 9D", "Angus 8B")
names(supp.labs) <- c("1B", "4D","2D","9D","8B")

# Stacked Percent
hist_fam_sample = taxa_meta %>% 
  mutate(Family = fct_relevel(Family,"Ruminococcaceae", "Rikenellaceae","Prevotellaceae","Lachnospiraceae", "Bacteroidaceae", "Muribaculaceae","Christensenellaceae","Spirochaetaceae","Bacteroidales.RF16.group","Bacteroidales.uncultured","Mollicutes.RF39.uncultured", "Minorities")) %>%
  ggplot(aes(x=ID, y=value,fill=Family)) + 
  geom_bar(stat="identity", position="fill") +
  facet_wrap(vars(Farm),scales = "free_x",nrow=1,labeller = labeller(Farm = supp.labs)) +
  scale_color_brewer(palette="Set3") + 
  theme(plot.subtitle = element_text(vjust = 1), plot.caption = element_text(vjust = 1)) +
  labs(x = "Farm", y = "Relative abundance", fill = "Family") + 
  theme_bw() +  
  scale_fill_brewer(palette="Paired") + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.text.x = element_blank(), axis.ticks = element_blank(), legend.position = "none", axis.title.x = element_blank())
hist_fam_sample
```

#Forage percentage histogram
```{r}
forage_hist = taxa_meta %>%
 # dplyr::filter(Family == "Ruminococcaceae") %>%
  ggplot(aes(x=ID, y=Pforage)) +
  geom_bar(stat="identity", position="dodge", fill = '#BBCC33') +
  facet_wrap(vars(Farm),scales = "free_x",nrow=1) +
 # geom_jitter(alpha=0.5) +
 # geom_smooth() +
  labs(y = "Forage %") + 
  theme_bw() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")  +
 theme(axis.text.x = element_blank(), axis.ticks = element_blank(),axis.title.x=element_blank())
forage_hist
```

#STEC percentage histogram
```{r}
STEC_hist = taxa_meta %>%
 # dplyr::filter(Family == "Ruminococcaceae") %>%
  ggplot(aes(x=ID, y=STEC)) +
  geom_bar(stat="identity", position="dodge", fill = 'red') +
  facet_wrap(vars(Farm),scales = "free_x",nrow=1) +
 # geom_jitter(alpha=0.5) +
 # geom_smooth() +
  labs(y = "STEC") + 
  theme_bw() +  
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")  +
 theme(axis.text.x = element_blank(), axis.ticks = element_blank(),axis.title.x=element_blank(),axis.text = element_blank())
STEC_hist
```

```{r}
hist_forage_sam = ggarrange(hist_fam_sample, forage_hist, STEC_hist, nrow = 3, ncol = 1, align="hv",labels = c("C","D","E"), heights = c(1.5,.5,.5))
hist_forage_sam
```
#Arranging histograms as a final figure
```{r}
Taxa_gut_sam = ggarrange(Taxa_gut, hist_forage_sam, nrow = 2, ncol = 1, align="hv", heights = c(1,1.5))
Taxa_gut_sam
```
#SAVING PLOTS AS A PDF 
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2020/june")

# mm to inch
setWidth = 183*0.039370

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='histogram_farm_STEC.pdf',width=setWidth,height=9,pointsize=setFontSize)
Taxa_gut_sam
dev.off()

```
