---
title: "Longitudinal_cattle_microbiome"
author: "Karla Vasco"
date: "5/5/2020"
output: html_document
---

```{r}
library("phyloseq")
library("ggplot2")
library("ape")
library("gridExtra")
library("ggpubr")
library("breakaway")
library("ggsignif")
library(forcats)

```

```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in otu table
otu_table = read.csv("otu_table_97.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table,rownames=TRUE)
#read in taxonomy
taxonomy = read.csv("taxonomy_97.csv",sep=",",row.names=1)
taxonomy=as.matrix(taxonomy)
#read in metadata
metadata=read.table("metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
#read in tree
phy_tree = read_tree("tree_97.nwk")
```

```{r}
#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
META$Phase=as.factor(META$Phase)
META$STEC=as.factor(META$STEC)

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)
taxa_names(phy_tree)

# Same sample names
sample_names(OTU)
sample_names(META)

#(tree was already imported as a phyloseq object)
physeq_zero=phyloseq(OTU,TAX,META,phy_tree)
physeq_zero
```

#Sorting physeq element 
```{r}
physeq_hindgut= subset_samples(physeq_zero, Type%in%c("RAJ","FG"))
physeq_long= subset_samples(physeq_hindgut, Longitudinal%in%c("Yes"))    
 

#physeq = subset_samples(physeq_long, Farm%in%c("1B"))
physeq = physeq_long
physeq = subset_samples(physeq_long, STEC%in%c("0","1"))

```

#Rarefraction
```{r}
physeq_rarefied <- rarefy_even_depth(physeq, sample.size = min(sample_sums(physeq)),
  rngseed = FALSE, replace = TRUE, trimOTUs = TRUE, verbose = TRUE)
```

#Taxa data frame
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots")
glom_fam <- tax_glom(physeq, taxrank = 'Family')
data_fam <- psmelt(glom_fam) # create dataframe from phyloseq object
write.csv(data_fam, file = "longitudinaL_family_ab.csv")
```

#LEFSE file
```{r}
library(metagMisc)
#physeq2 = filter_taxa(physeq, function(x) mean(x) > 0.1, TRUE)
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots")
physeq = subset_samples(physeq_long, Farm%in%c("8B"))
glom_gen <- tax_glom(physeq, taxrank = 'Genus')
cattle_df_gen = phyloseq_to_df(glom_gen, addtax = T, addtot = F, addmaxrank = F, sorting = "Abundance")
cattle_df_gen = cattle_df_gen[,c(-1,-8)]

cattle_df_gen2 = tidyr::unite(cattle_df_gen, col = "Taxa", Kingdom:Genus, sep = "|", remove = TRUE, na.rm = FALSE)

cattle_df_gen3 = as.data.frame(t(cattle_df_gen2))
names(cattle_df_gen3) <- as.matrix(cattle_df_gen3[1, ])
cattle_df_gen3 <- cattle_df_gen3[-1, ]
cattle_df_gen3[] <- lapply(cattle_df_gen3, function(x) type.convert(as.character(x)))
cattle_df_gen3


cattle_df_gen4 <- merge(cattle_df_gen3,metadata,by="row.names", all=F)
lefse_stec_epec_pos = cattle_df_gen4 %>%  
  dplyr::select(1,Phase,2:362)

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots")
write.table(lefse_stec_epec_pos, file = "longitudinal_8B.txt", sep = '\t', col.names = TRUE, row.names = F)
```

#Alpha diversity
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq, measures = c("Shannon", "Chao1"))
df_alpha2 <- data.frame(alpha_diversity, sample_data(physeq))
df_alpha = df_alpha2 %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))
df <- reshape2::melt(df_alpha, measure.var=c("Shannon","Chao1"),id.vars=c("Phase","Farm","STEC","STEC_presence"))
df$value = as.numeric(df$value)
df$Phase =as.factor(df$Phase)

#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df, variable == "Shannon")
chao = dplyr::filter(df, variable == "Chao1")
```

```{r}
alpha_diversity = estimate_richness(physeq, measures = c("Shannon","Chao1"))
alpha.stats <- cbind(alpha_diversity, sample_data(physeq))

F1B = alpha.stats %>%
dplyr::filter(Farm == "1B")
kruskal.test(Shannon~Phase, data = F1B)

F2D = alpha.stats %>%
dplyr::filter(Farm == "2D")
kruskal.test(Shannon~Phase, data = F2D)

F4D = alpha.stats %>%
dplyr::filter(Farm == "4D")
kruskal.test(Shannon~Phase, data = F4D)

F8B = alpha.stats %>%
dplyr::filter(Farm == "8B")
kruskal.test(Shannon~Phase, data = F8B)

F9D = alpha.stats %>%
dplyr::filter(Farm == "9D")
kruskal.test(Shannon~Phase, data = F9D)

F1Bc = alpha.stats %>%
dplyr::filter(Farm == "1B")
kruskal.test(Chao1~Phase, data = F1B)

F2Dc = alpha.stats %>%
dplyr::filter(Farm == "2D")
kruskal.test(Chao1~Phase, data = F2D)

F4Dc = alpha.stats %>%
dplyr::filter(Farm == "4D")
kruskal.test(Chao1~Phase, data = F4D)

F8Bc = alpha.stats %>%
dplyr::filter(Farm == "8B")
kruskal.test(Chao1~Phase, data = F8B)

F9Dc = alpha.stats %>%
dplyr::filter(Farm == "9D")
kruskal.test(Chao1~Phase, data = F9D)

```


```{r}
################ Modify these variables as needed #################
x_axis = df_alpha$Phase #variable to test
x_axis_sha = shannon$Phase #variable to test
x_axis_chao = chao$Phase #variable to test
my_comparisons <-  list(c("1", "2"), c("2", "3"), c("3", "4"), c("1", "4"))#list(c("2", "3"), c("3", "4"), c("2", "4"))
#p-value between groups 
stat_lab_position_x = 0.3
```
#ALPHA FARM COLORS
```{r}
##################### Running box plot #############################
supp.labs <- c("1B   p=0.77", "2D   p = 0.41", "4D   p=0.02", "9D   p=0.11", "8B  p=2.7e-06")
names(supp.labs) <- c("1B", "2D","4D","9D","8B")

supp.labs2 <- c("1B   p=0.0008", "2D  p = 0.005", "4D   p=0.9", "9D   p=0.02", "8B   p=7.33e-08")
names(supp.labs2) <- c("1B", "2D","4D","9D","8B")

shannon_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Shannon, fill = Farm, color = STEC.shedding, shape = STEC.shedding)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs))+
  geom_point(alpha=0.5, position = position_jitterdodge()) +
  labs(y="Shannon", x="Phase",color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00")) +
  scale_color_manual(values =c("red","red","red","blue","red")) +  scale_shape_manual(values = c(24,25,23,20,3))
 # geom_text(aes(label=Study_ID),hjust=0, vjust=0)

chao1_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Chao1, fill = Farm, color = STEC.shedding,shape = STEC.shedding)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs2))+
  geom_point(alpha=0.5, position = position_jitterdodge()) +
  labs(y="Chao1", x="Phase", color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00")) +
  scale_color_manual(values =c("red","red","red","blue","red")) +  scale_shape_manual(values = c(24,25,23,20,3))

shannon_box
chao1_box
```

#ALPHA diversity, colors by positive and negative and shape by shedding
```{r}
##################### Running box plot #############################
supp.labs <- c("1B   p=0.77", "2D   p = 0.41", "4D   p=0.02", "9D   p=0.11", "8B  p=2.7e-06")
names(supp.labs) <- c("1B", "2D","4D","9D","8B")

supp.labs2 <- c("1B   p=0.0008", "2D  p = 0.005", "4D   p=0.9", "9D   p=0.02", "8B   p=7.33e-08")
names(supp.labs2) <- c("1B", "2D","4D","9D","8B")

shannon_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Shannon, fill = STEC, shape = STEC.shedding)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs))+
  geom_point(alpha=0.5, position = position_jitterdodge(), size = 3) +
  labs(y="Shannon", x="Phase",color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#56B4E9","#D55E00")) +  
  scale_shape_manual(values = c(24,25,23,20,3))
 # geom_text(aes(label=Study_ID),hjust=0, vjust=0)

chao1_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Chao1, fill = STEC, shape = STEC.shedding)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs2))+
  geom_point(alpha=0.5, position = position_jitterdodge(), size = 3) +
  labs(y="Chao1", x="Phase", color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#56B4E9","#D55E00")) +  
  scale_shape_manual(values = c(24,25,23,20,3))

shannon_box
chao1_box
```

#ALPHA diversity, colors and shape by positive and negative 
```{r}
##################### Running box plot #############################
supp.labs <- c("1B   p=0.66", "2D   p = 0.20", "4D   p=0.002", "9D   p=0.30", "8B  p=1.65e-05")
names(supp.labs) <- c("1B", "2D","4D","9D","8B")

supp.labs2 <- c("1B   p=0.018", "2D  p = 0.006", "4D   p=0.35", "9D   p=0.13", "8B   p=1.049e-07")
names(supp.labs2) <- c("1B", "2D","4D","9D","8B")

shannon_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Shannon, fill = STEC_presence, shape = STEC_presence)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs))+
  geom_point(alpha=0.5, position = position_jitterdodge(), size = 3) +
  labs(y="Shannon", x="Phase",color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#56B4E9","#D55E00")) +  
  scale_shape_manual(values = c(20,3))
 # geom_text(aes(label=Study_ID),hjust=0, vjust=0)

chao1_box <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Chao1, fill = STEC_presence, shape = STEC_presence)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs2))+
  geom_point(alpha=0.5, position = position_jitterdodge(), size = 3) +
  labs(y="Chao1", x="Phase", color = "STEC shedding", fill = "STEC shedding") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "none",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#56B4E9","#D55E00")) +  
  scale_shape_manual(values = c(20,3))

shannon_box
chao1_box
```
#legend alpha
```{r}
alpha_leg <- df_alpha %>% #shannon_box, chao1_box
  mutate(x_axis = fct_relevel(Phase, "1","2","3","4")) %>% #Optional if you need a specific x_axis order
  ggplot(aes(x = Phase, y = Chao1, fill = STEC_presence, shape = STEC_presence)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  facet_wrap(vars(Farm),nrow=1,ncol=5,scales="free", labeller = labeller(Farm = supp.labs2))+
  geom_point(alpha=0.5, position = position_jitterdodge(), size = 3) +
  labs(y="Chao1", x="Phase", color = "STEC", fill = "STEC", shape = "STEC") + 
    theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(legend.position = "top",axis.title.x=element_blank()) + 
 # stat_compare_means(comparisons = my_comparisons, step.increase = .15, symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) + 
  scale_fill_manual(values =c("#56B4E9","#D55E00")) +  
  scale_shape_manual(values = c(20,3))
# Extract the legend. Returns a gtable
leg_alpha <- get_legend(alpha_leg)

# Convert to a ggplot and print
legend_alpha_STEC = as_ggplot(leg_alpha)
legend_alpha_STEC
```

#BETA DIVERSITY

#Stats beta

#1B beta
#Stats beta
```{r}
physeq = subset_samples(physeq_rarefied, Farm%in%c("1B"))

## PERMANOVA
library(vegan)
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Phase)
adonis_wunifrac

wu_plot = plot_ordination(physeq, beta_wu, color="Type", shape="Phase") + theme_bw()+labs(colour = "Farm", shape="Phase")
wu_plot
```

```{r}
physeq = subset_samples(physeq_long, Farm%in%c("1B"))
beta_wu <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time

supp.labs <- c("1B  R^2=0.1, p=0.001")
names(supp.labs) <- c("1B")

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
beta_meta$STEC = as.factor(beta_meta$STEC)
beta_meta = beta_meta %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))

wu_1BL = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", shape = "STEC_presence", palette = c("#E69F00","#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (16.5%)", y = "PCoA 2 (5.1%)", colour = "Phase", fill = "Phase", shape = "Phase") + 
  rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") +
  scale_shape_manual(values = c(20,3))#annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 
wu_1BL
```

#2D beta
```{r}
physeq = subset_samples(physeq_rarefied, Farm%in%c("2D"))

## PERMANOVA
library(vegan)
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Phase)
adonis_wunifrac

beta_wu <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
wu_plot = plot_ordination(physeq, beta_wu, color="STEC", shape="Phase") +  theme_bw()+labs(shape="Phase")
wu_plot
```

```{r}
beta_wu <- ordinate(physeq, "PCoA", "wunifrac")
supp.labs <- c("2D R^2=0.08, p=0.001")
names(supp.labs) <- c("2D")

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
beta_meta$STEC = as.factor(beta_meta$STEC)
beta_meta = beta_meta %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))
wu_2DL = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", shape = "STEC_presence", palette = c("#E69F00","#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (6.8%)", y = "PCoA 2 (5.9%)", colour = "Phase", fill = "Phase", shape = "Phase") + 
  rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") +   scale_shape_manual(values = c(20,3))
#+ geom_text(aes(label=Study_ID),hjust=0, vjust=0)#annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 
wu_2DL
```

#4D beta
```{r}
physeq = subset_samples(physeq_rarefied, Farm%in%c("4D"))

## PERMANOVA
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Phase)
adonis_wunifrac

beta_wu <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
wu_plot = plot_ordination(physeq, beta_wu, color="STEC", shape="Phase") +  theme_bw()+labs(shape="Phase")
wu_plot
```

```{r}
physeq = subset_samples(physeq_long, Farm%in%c("4D"))
beta_wu <- ordinate(physeq, "PCoA", "wunifrac")
supp.labs <- c("4D  R^2=0.045, p=0.001")
names(supp.labs) <- c("4D")

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
beta_meta$STEC = as.factor(beta_meta$STEC)
beta_meta = beta_meta %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))
wu_4DL = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", shape = "STEC_presence", palette = c("#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (11.1%)", y = "PCoA 2 (6.6%)", colour = "Phase", fill = "Phase", shape = "Phase") + 
  rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none")+
  scale_shape_manual(values = c(20,3,20))   #annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 
wu_4DL
```
#8B beta
```{r}
physeq = subset_samples(physeq_rarefied, Farm%in%c("8B"))

## PERMANOVA
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Phase)
adonis_wunifrac

beta_wu <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
wu_plot = plot_ordination(physeq, beta_wu, color="STEC", shape="Phase") +  theme_bw()+labs(shape="Phase")
wu_plot
```

```{r}

physeq = subset_samples(physeq_long, Farm%in%c("8B"))
beta_wu <- ordinate(physeq, "PCoA", "wunifrac")
supp.labs <- c("8B  R^2=0.15, p=0.001")
names(supp.labs) <- c("8B")

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
beta_meta$STEC = as.factor(beta_meta$STEC)
beta_meta = beta_meta %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))
wu_8B = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", fill = "Phase", shape = "STEC_presence", palette = c("#E69F00","#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F , mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (10.1%)", y = "PCoA 2 (9.1%)", colour = "Phase", fill = "Phase", shape = "Phase") + 
 # rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") + scale_shape_manual (values = c(20,3)) #annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 
wu_8B
```
#9D beta
```{r}
physeq = subset_samples(physeq_rarefied, Farm%in%c("9D"))

## PERMANOVA
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Phase)
adonis_wunifrac

beta_wu <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
wu_plot = plot_ordination(physeq, beta_wu, color="STEC", shape="Phase") +  theme_bw()+labs(shape="Phase")
wu_plot
```

```{r}
supp.labs <- c("9D  R^2=0.105, p=0.013")
names(supp.labs) <- c("9D")

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
beta_meta$STEC = as.factor(beta_meta$STEC)
beta_meta = beta_meta %>% 
  mutate(STEC_presence = dplyr::case_when(STEC == "0" ~ "Negative", T ~ "Positive"))
wu_9Dt = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", shape = "STEC_presence", fill = "Phase",palette = c("#E69F00","#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (13.6%)", y = "PCoA 2 (8.5%)", colour = "Phase", fill = "Phase", shape = "Phase") + 
 # rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "none") +  scale_shape_manual(values = c(20,3)) #annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 
wu_9Dt
```

#BETA legend
```{r}
plot_leg_WU = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Phase", shape = "STEC_presence", fill = "Phase",palette = c("#E69F00","#D55E00","#56B4E9","#0072B2"), ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +
  labs(x = "PCoA 1 (17.9%)", y = "PCoA 2 (12.7%)", colour = "Phase", fill = "Phase", shape = "STEC") + 
 # rremove("legend") + 
  facet_wrap(vars(Farm),nrow=1,ncol=1,scales="free",labeller = labeller(Farm = supp.labs))+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.position = "top") +  scale_shape_manual(values = c(20,3)) #annotate("text", x = -0.12, y = -0.08, label = "PERMANOVA, p <0.001\nANOVA, p=0.045", colour = "black") 

leg_wuL <- get_legend(plot_leg_WU)

# Convert to a ggplot and print
legend_wuL = as_ggplot(leg_wuL)
legend_wuL
```


#Diversity plot NEW
```{r}
beta_div_long = ggarrange(wu_1BL,wu_2DL,wu_4DL,wu_8B,wu_9Dt, nrow = 1,ncol=5,align="hv")
long_div = ggarrange(legend_alpha_STEC, shannon_box, chao1_box, legend_wuL, beta_div_long, nrow = 5, ncol = 1, align="hv", heights = c(.3,1,1,.3, 1.5), labels = c("","A","B","","C"))
long_div
```

```{r}

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021")
# mm to inch
setWidth = 12

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='Longitudinal_diversity_rarefied.pdf',width=setWidth,height=7,pointsize=setFontSize)
long_div
dev.off()

```
#Filtering samples for LEFSE
```{r}
physeq_two = subset_taxa(physeq, Order != "Chloroplast")
physeq_three = subset_taxa(physeq_two, Family != "Mitochondria")
physeq_hindgut <- subset_samples(physeq_three, Type%in%c("RAJ","FG"))
physeq_long= subset_samples(physeq_hindgut, Longitudinal%in%c("Yes"))    
 physeq = subset_samples(physeq_long, STEC%in%c("0","1"))
```

#LEFSE file
```{r}
library(metagMisc)
#physeq2 = filter_taxa(physeq, function(x) mean(x) > 0.1, TRUE)
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots")
physeq3 = transform_sample_counts(physeq, function(x) x / sum(x) ) #Normalizing counts
glom_gen <- tax_glom(physeq3, taxrank = 'Genus') #Merging OTU to a genus level
cattle_df_gen = phyloseq_to_df(glom_gen, addtax = T, addtot = F, addmaxrank = F, sorting = "Abundance") #Converting to data frame
cattle_df_gen = cattle_df_gen[,c(-1,-8)] #eliminating column numbers

cattle_df_gen2 = tidyr::unite(cattle_df_gen, col = "Taxa", Kingdom:Genus, sep = "|", remove = TRUE, na.rm = FALSE) #Merging taxa info to a LEFSE format

cattle_df_gen3 = as.data.frame(t(cattle_df_gen2)) 
names(cattle_df_gen3) <- as.matrix(cattle_df_gen3[1, ])
cattle_df_gen3 <- cattle_df_gen3[-1, ]
cattle_df_gen3[] <- lapply(cattle_df_gen3, function(x) type.convert(as.character(x)))
cattle_df_gen3

cattle_df_gen_4 <- merge(cattle_df_gen3,metadata,by="row.names", all=T)

lefse_stec_9d_long = cattle_df_gen_4 %>% dplyr::filter(Farm%in%c("9D")) %>% dplyr::filter(Type%in%c("FG", "RAJ")) %>% 
  dplyr::select(1,STEC,Farm,2:362)

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/tables_plots")
write.table(lefse_stec_9d_long, file = "lefse_stec_9d_long.txt", sep = '\t', col.names = TRUE, row.names = F)

#Modify in Excel Row.names for ID (optional)
#Then submit the LEFSE analysis here http://huttenhower.org/galaxy/
```
