---
title: "R Notebook"
output: html_notebook
---
#Libraries
```{r}
library("phyloseq")
library("ggplot2")
library("ape")
library("gridExtra")
library("ggpubr")
library("ggsignif")
library(forcats)
library(vegan)
library(viridis)
library(cowplot)
```

#Importing files for microbiome analysis with phyloseq
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/data/Microbiome_2019/97_table_data/phyloseq_objects")

#read in otu table
otu_table = read.csv("otu_table_97.csv",sep=",",row.names=1)
otu_table=as.matrix(otu_table,rownames=TRUE)
#read in taxonomy
taxonomy = read.csv("taxonomy_97.csv",sep=",",row.names=1)
taxonomy=as.matrix(taxonomy)
#read in metadata
metadata=read.table("metadata_up.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree_97.nwk")
```

##Generating a phyloseq object 
```{r}
#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)
META$Phase=as.factor(META$Phase)
META$STEC=as.factor(META$STEC)
META$STEC=as.factor(META$STEC)
META$STEC_prev_group <- factor(META$STEC_prev_group, levels = c("Low","High"))

# Sanity checks for consistent OTU names
taxa_names(TAX)
taxa_names(OTU)
taxa_names(phy_tree)

# Same sample names
sample_names(OTU)
sample_names(META)

#(tree was already imported as a phyloseq object)
physeq=phyloseq(OTU,TAX,META,phy_tree)
```

#Filtering samples and taxa for further analyses
```{r}
physeq_two = subset_taxa(physeq, Order != "Chloroplast")
physeq_three = subset_taxa(physeq_two, Family != "Mitochondria")
physeq_hindgut <- subset_samples(physeq_three, Type%in%c("RAJ","FG"))
physeq_STEC <- subset_samples(physeq_hindgut, STEC%in%c("1","0"))
physeq = physeq_STEC
```

#Alpha-diversity calculation  (Shannon and Chao1 indices) 
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq, measures = c("Shannon", "Chao1"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq))
df <- reshape2::melt(df_alpha, measure.var=c("Shannon","Chao1"),id.vars=c("Phase","Farm","STEC","Type","LACT","Lymph","Neut","Monocytes","Lymphocyte.Monocyte.ratio","Lymphocytes","STEC_prev_group", "STEC_comp","STEC_EPEC_comp"))
df$value = as.numeric(df$value)
df$Phase =as.factor(df$Phase)

#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df, variable == "Shannon")
chao = dplyr::filter(df, variable == "Chao1")
```


#Alpha-diversity comparisons between farms (plot and stats)
```{r}
library(ggsci)
my_comparisons <- list(c("1B", "2D"), c("2D", "4D"), c("4D", "9D"), c("9D","8B"), c("2D","9D"), c("1B","8B"))

##################### Running box plot #############################

sha_farm <- shannon %>% #shannon_box, chao1_box
  mutate(Farm = fct_relevel(Farm, "1B","2D","4D","9D","8B")) %>% #Optional if you need a specific x_axis order 
  ggplot(aes(x = Farm, y = value, fill = STEC_prev_group)) +
 # geom_violin(width=1.4) +
  geom_boxplot(outlier.shape = NA) +
  #geom_boxplot(outlier.shape = NA) +
 # geom_point(alpha=0.5, position = position_jitterdodge()) +
  geom_jitter(color="black", size=0.4, alpha=0.5) +
  labs(y="Shannon", x="Farm") +
    annotate("text", x = 3, y = 7.8, label = "Kruskal-Wallis, p < 2.2e-16", colour = "black") +
  ylim(3,8) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(legend.position = "none",axis.title.x=element_blank()) + stat_compare_means(comparisons = my_comparisons, label.y = c(6.1,6.3,6.5,6.65,6.8,7.2), symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = F) +
  #stat_compare_means(label.y = 7.7, label.x.npc = .3) +
   scale_fill_aaas()
#  scale_fill_manual(values =c("#73D055FF","#404788FF","#56B4E9","#E69F00","#D55E00")) + scale_shape_manual(values = c(20,3,20))

cha_farm <- chao %>% #shannon_box, chao1_box
 mutate(Farm = fct_relevel(Farm, "1B","2D","4D","9D","8B")) %>% #Optional if you need a specific x_axis order 
  ggplot(aes(x = Farm, y = value, fill = STEC_prev_group)) +
 # geom_violin(width=1.4) +
  geom_boxplot(outlier.shape = NA) +
#  geom_point(alpha=0.5, position = position_jitterdodge()) +
  geom_jitter(color="black", size=0.4, alpha=0.5) +
   labs(y="Chao1", x="Farm") + 
    annotate("text", x = 3, y = 2400, label = "Kruskal-Wallis, p < 2.2e-16", colour = "black") +
  ylim(100,2500) +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(legend.position = "none",axis.title.x=element_blank()) + stat_compare_means(comparisons = my_comparisons, label.y = c(1500,1600,1700,1800,1900,2100), symnum.args = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1), symbols = c("****", "***", "**", "*", "ns")), tip.length=.02, hide.ns = FALSE) +
#  stat_compare_means(label.y = 2400, label.x.npc = .3) +
 scale_fill_aaas()
  # scale_fill_manual(values =c("#73D055FF","#404788FF","#56B4E9","#E69F00","#D55E00")) +scale_shape_manual(values=c(20,3,20))
sha_farm
cha_farm
```

##Arrange alpha-diversity plots as a single figure 
```{r}
alphaF = ggarrange(sha_farm,cha_farm,nrow = 1, ncol = 2, align="hv", labels = c("A","B")) 
```

##Export alpha-diversity plot as a PDF
```{r}

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2020/june")
# mm to inch
setWidth = 183*0.039370

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='alpha_div_farm.pdf',width=setWidth,height=4.5,pointsize=setFontSize)
alphaF
dev.off()

```

#Alpha diversity plot between STEC prevalence groups
```{r}
library(tidyverse)
library("ggsci") #Color palettes
library(ggpubr)
library(rstatix)
library(scales)
```

##Shannon index plot
```{r}
shannon_boxplot_prev = ggboxplot(shannon, x = "STEC_prev_group", y = "value", color = "STEC_prev_group", palette = "lancet", fill = "STEC_prev_group", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F, legend = "none") +labs(y = "Shannon", colour = "STEC prevalence", fill = "STEC prevalence") + scale_fill_lancet(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 6.3)

shannon_boxplot_prev
```
##Chao1 index plot
```{r}
chao1_boxplot_prev = ggboxplot(chao, x = "STEC_prev_group", y = "value", color = "STEC_prev_group", palette = "lancet", fill = "STEC_prev_group", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F, legend = "none") +labs(y = "Chao1", colour = "STEC prevalence", fill = "STEC prevalence") + scale_fill_lancet(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 6.3)

chao1_boxplot_prev
```

##Arrange alpha-diversity plots as a single figure 
```{r}
alpha_STEC_groups = ggarrange(shannon_boxplot_prev,chao1_boxplot_prev,nrow = 2, ncol = 1, align="hv", labels = c("A","B")) 
```


#Weighted unifrac 
##Weighted Unifrac ordination and preliminary plot
```{r}
beta_wu_farm <- ordinate(physeq, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
beta_wu_plot = plot_ordination(physeq, beta_wu_farm, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```

##Weighted Unifrac stats among farms
```{r}
## PERMANOVA
library(vegan)
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Farm)
adonis_wunifrac

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$Farm, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")
```

##Weighted Unifrac plot among farms
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.23, ", paste(italic('p')),"=0.001"))

beta_scatter = as.data.frame(beta_wu_farm[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_wuF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Farm", shape = "STEC_pos", palette =   c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00"),ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (18.6%) ", y = "PCoA 2 (6.3%)", colour = "STEC_prev_group", fill = "Farm") + rremove("legend") +annotate("text", x = -0.08, y = -0.08, label = label_perm, colour = "black") +scale_shape_manual(values=c(20,3))
pmain_wuF

xdens <- axis_canvas(pmain_wuF, axis = "x")+
  geom_density(data = beta_meta, aes(x = Axis.1, fill = Farm),
              alpha = 0.7, size = 0.2)+
   theme(legend.position = "none") +
  scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00"))
# Marginal densities along y axis
# Need to set coord_flip = TRUE, if you plan to use coord_flip()
ydens <- axis_canvas(pmain_wuF, axis = "y", coord_flip = TRUE)+
     theme(legend.position = "none") +
  geom_density(data = beta_meta, aes(x = Axis.2, fill = Farm),
                alpha = 0.7, size = 0.2)+
  coord_flip()+
  scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00"))
p1 <- insert_xaxis_grob(pmain_wuF, xdens, grid::unit(.2, "null"), position = "top")
p2<- insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = "right")

wu_plotF = ggdraw(p2)
wu_plotF
```

#Bray-Curtis dissimilarity
##Bray-Curtis dissimilarity ordination and preliminary plot
```{r}
beta_bray_farm <- ordinate(physeq, "PCoA", "bray") #RUN this only ONCE because it takes a lot of time
beta_bray_plot = plot_ordination(physeq, beta_bray_farm, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_bray_plot
```

#Bray-Curtis dissimilarity statistics betweeen STEC prevalence groups
```{r}
## PERMANOVA
library(vegan)
bray_dist = phyloseq::distance(physeq, method="bray") #RUN this only once because it takes a lot of time
adonis_bray = adonis(bray_dist ~ sample_data(physeq)$STEC_prev_group)
adonis_bray

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
bray_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$STEC_prev_group, type=c("median"))
anova(bray_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(bray_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(bray_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(bray_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(bray_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(bray_disp)

scores(bray_disp, 1:4, display = "centroids")

```

#Bray-Curtis dissimilarity plot between STEC prevalence groups
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.07, ", paste(italic('p')),"=0.001"))

beta_scatter = as.data.frame(beta_bray_farm[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_brayF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "STEC_prev_group", shape = "STEC_pos", palette = c("#EE0000FF","#3B4992FF"),ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (18.8%) ", y = "PCoA 2 (7.2%)", colour = "STEC prevalence", fill = "STEC prevalence", shape = "STEC")  +annotate("text", x = -0.3, y = -0.23, label = label_perm, colour = "black")+ #+annotate("text", x = -0.3, y = -0.3, label = label_anosim, colour = "black") + rremove("legend")
scale_shape_manual(values=c(20,3))
pmain_brayF
#pallete = c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00")
xdens <- axis_canvas(pmain_brayF, axis = "x")+
  geom_density(data = beta_meta, aes(x = Axis.1, fill = STEC_prev_group),
              alpha = 0.7, size = 0.2)+
   theme(legend.position = "none") +
  #scale_fill_aaas()
scale_fill_manual(values=c("#EE0000FF","#3B4992FF"))
# scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00"))
# Marginal densities along y axis
# Need to set coord_flip = TRUE, if you plan to use coord_flip()
ydens <- axis_canvas(pmain_brayF, axis = "y", coord_flip = TRUE)+
     theme(legend.position = "none") +
  geom_density(data = beta_meta, aes(x = Axis.2, fill = STEC_prev_group),
                alpha = 0.7, size = 0.2)+
  coord_flip()+
 #  scale_fill_aaas()
scale_fill_manual(values=c("#EE0000FF","#3B4992FF"))
# scale_fill_manual(values =c("#73D055FF","#56B4E9","#404788FF","#D55E00","#E69F00"))
p1 <- insert_xaxis_grob(pmain_brayF, xdens, grid::unit(.2, "null"), position = "top")
p2<- insert_yaxis_grob(p1, ydens, grid::unit(.2, "null"), position = "right")

bray_plotF = ggdraw(p2)
bray_plotF
```

##Arrange alpha- and beta-diversity plots as a single figure 
```{r}
beta_div = ggarrange(wu_plotF,bray_plotF,nrow = 2, ncol = 1, align="hv", labels = c("C","D")) 
diversity = ggarrange(alpha_STEC_groups,beta_div,nrow = 1, ncol = 2, align="hv", widths = c(1,1.5)) 
```

##Export alpha-diversity plot as a PDF
```{r}

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021")
# mm to inch
setWidth = 10

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='diversity_plots.pdf',width=setWidth,height=9,pointsize=setFontSize)
diversity
dev.off()
```


#Alpha diverstiy STEC shedders

##stx+eae- shedders vs. controls

###Chao1 plot STEC vs controls
```{r}
chao_STEC_comp = chao%>%
  dplyr::filter(STEC_comp != "NA")

chao_boxplot_stec = ggboxplot(chao_STEC_comp, x = "STEC_comp", y = "value", color = "STEC", palette = "lancet", fill = "STEC", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F,legend = "none") +labs(y = "Chao1", colour = "STEC", fill = "STEC") + scale_fill_lancet(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 1800)

chao_boxplot_stec
```

###Shannon plot STEC vs controls
```{r}
shannon_STEC_comp = shannon%>%
  dplyr::filter(STEC_comp != "NA")

shannon_boxplot_stec = ggboxplot(shannon_STEC_comp, x = "STEC_comp", y = "value", color = "STEC", palette = "lancet", fill = "STEC", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F, legend = "none") +labs(y = "Shannon", colour = "STEC", fill = "STEC") + scale_fill_lancet(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 6.3)

shannon_boxplot_stec
```

##stx+eae+ shedders vs. controls

###Chao1 plot EHEC vs controls
```{r}
chao_STEC_EPEC_comp = chao%>%
  dplyr::filter(STEC_EPEC_comp != "NA")

chao_boxplot_epec = ggboxplot(chao_STEC_EPEC_comp, x = "STEC_EPEC_comp", y = "value", color = "STEC", palette = "jco", fill = "STEC", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F,legend = "none") +labs(y = "Chao1", colour = "STEC", fill = "STEC") + scale_fill_jco(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 1800)

chao_boxplot_epec
```

###Shannon plot EHEC vs controls
```{r}
shannon_STEC_EPEC_comp = shannon%>%
  dplyr::filter(STEC_EPEC_comp != "NA")

shannon_boxplot_epec = ggboxplot(shannon_STEC_EPEC_comp, x = "STEC_EPEC_comp", y = "value", color = "STEC", palette = "jco", fill = "STEC", add = c("jitter"), notch = F, outlier.shape = NA, xlab = F,legend = "none") +labs(y = "Shannon", colour = "STEC", fill = "STEC") + scale_fill_jco(alpha = 0.5) +
  stat_compare_means(label.x.npc = c(.25),label.y = 6.4)

shannon_boxplot_epec
```

```{r fig.height=3, fig.width=10}
stec_boxplots <- ggarrange(shannon_boxplot_stec,chao_boxplot_stec,ncol = 2, nrow = 1, align = "hv", labels = c("A","B"))
stec_boxplots
stec_epec_boxplots <- ggarrange(shannon_boxplot_epec,chao_boxplot_epec,ncol = 2, nrow = 1, align = "hv", labels = c("C","D"))
stec_epec_boxplots

alpha_stec <-  ggarrange(stec_boxplots,stec_epec_boxplots,ncol = 2, nrow = 1, align = "hv")
alpha_stec
```


#Weighted Unifrac stats between STEC shedders and controls
```{r}
beta_wu_STEC <- ordinate(physeq_STEC, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
beta_wu_plot = plot_ordination(physeq_STEC, beta_wu_STEC, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```

```{r}
#WU stats
## PERMANOVA
library(vegan)
physeq_STEC <- subset_samples(physeq, STEC_comp%in%c("Control","STEC+EPEC-"))

wuinfrac_dist = phyloseq::distance(physeq_STEC, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq_STEC)$STEC_comp)
adonis_wunifrac

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq_STEC)$STEC_comp, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")

```

```{r}
#WU stats
## PERMANOVA
library(vegan)
physeq_STEC_EPEC <- subset_samples(physeq, STEC_EPEC_comp%in%c("Control","STEC+EPEC+"))

wuinfrac_dist = phyloseq::distance(physeq_STEC_EPEC, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq_STEC_EPEC)$STEC_EPEC_comp)
adonis_wunifrac

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq_STEC_EPEC)$STEC_EPEC_comp, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")

```


#Weighted unifrac ordination and preliminary plot of STEC shedders and controls
```{r}
beta_wu_STEC <- ordinate(physeq_STEC, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
beta_wu_plot = plot_ordination(physeq_STEC, beta_wu_STEC, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```

#Weighted Unifrac plot between STEC shedders and controls
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.08, ", paste(italic('p')),"=0.001"))
library(lemon)

beta_scatter = as.data.frame(beta_wu_STEC[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_wuF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "STEC_comp",ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F, legend = "top") +labs(x = "PCoA 1 (27.3%) ", y = "PCoA 2 (6.9%)", colour = "", fill = "") +annotate("text", x = -0.05, y = 0.04, label = label_perm, colour = "black") +scale_shape_manual(values=c(20,3)) + scale_fill_lancet(alpha = 0.5) +scale_color_lancet() 

pmain_wuF = reposition_legend(pmain_wuF,'top left')

```

```{r}
beta_wu_STEC <- ordinate(physeq_STEC, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
beta_wu_plot = plot_ordination(physeq_STEC, beta_wu_STEC, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```

```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.08, ", paste(italic('p')),"=0.001"))
library(lemon)

beta_scatter = as.data.frame(beta_wu_STEC[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_wuF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "STEC_comp",ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F, legend = "top") +labs(x = "PCoA 1 (27.3%) ", y = "PCoA 2 (6.9%)", colour = "", fill = "") +annotate("text", x = -0.02, y = 0.04, label = label_perm, colour = "black") +scale_shape_manual(values=c(20,3)) + scale_fill_lancet(alpha = 0.5) +scale_color_lancet() 

pmain_wuF = reposition_legend(pmain_wuF,'top left')

```

#Weighted unifrac beta diversity
```{r}
beta_wu_STEC_EPEC <- ordinate(physeq_STEC_EPEC, "PCoA", "wunifrac") #RUN this only ONCE because it takes a lot of time
beta_wu_plot = plot_ordination(physeq_STEC_EPEC, beta_wu_STEC_EPEC, type="Farm", color="Farm", shape="Farm", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "Farm") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.02, ", paste(italic('p')),"=0.001"))
library(lemon)

beta_scatter = as.data.frame(beta_wu_STEC_EPEC[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_wuF_epec = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "STEC_EPEC_comp",ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F, legend = "top") +labs(x = "PCoA 1 (20.4%) ", y = "PCoA 2 (7.5%)", colour = "", fill = "") +annotate("text", x = -0.05, y = 0.07, label = label_perm, colour = "black") +scale_shape_manual(values=c(20,3)) + scale_fill_jco(alpha = 0.5) +scale_color_jco() 

pmain_wuF_epec = reposition_legend(pmain_wuF_epec,'top left')

```

```{r}
beta_STEC_EPEC = ggarrange(pmain_wuF, pmain_wuF_epec, nrow = 1, ncol = 2, align="hv", labels = c("E","F"))
beta_STEC_EPEC
stec_PLOTS = ggarrange(alpha_stec,beta_STEC_EPEC,nrow = 2, ncol = 1, align="hv") 
stec_PLOTS
```

```{r}

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/")
# mm to inch
setWidth = 12

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='2021_05_19_STECsamples_div.pdf',width=setWidth,height=6,pointsize=setFontSize)
stec_PLOTS
dev.off()

```


#Forage percentage beta-diversity comparisons 
#Weighted-Unifrac stats Forage
```{r}
## PERMANOVA
library(vegan)
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$Pforage)
adonis_wunifrac

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$Pforage, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")

```

##Weighted-Unifrac plot Forage
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.13, ", paste(italic('p')),"=0.001"))

beta_scatter = as.data.frame(beta_wu_farm[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_wuF_forage = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Pforage",ellipse = F, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (18.6%) ", y = "PCoA 2 (6.3%)", colour = "Forage %", fill = "Forage %") + scale_color_viridis(option="D") + annotate("text", x = -0.05, y = 0.18, label = label_perm, colour = "black")
pmain_wuF_forage
```

##Bray-Curtis dissimilarity stats Forage
```{r}
## PERMANOVA
library(vegan)
bray_dist = phyloseq::distance(physeq, method="bray") #RUN this only once because it takes a lot of time
adonis_bray = adonis(bray_dist ~ sample_data(physeq)$Pforage)
adonis_bray

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$Pforage, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")
```

##Bray-Curtis dissimilarity plot Forage
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.13, ", paste(italic('p')),"=0.001"))

beta_scatter = as.data.frame(beta_bray_farm[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
beta_meta$Phase = as.factor(beta_meta$Phase)
pmain_bray_forage = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "Pforage",ellipse = F, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (18.8%) ", y = "PCoA 2 (7.2%)", colour = "Forage %", fill = "Forage %") + scale_color_viridis(option="D") + annotate("text",  x = -0.25, y = 0.3, label = label_perm, colour = "black")
pmain_bray_forage
```

#Arrange beta-diversity plots of forage percentage in diet
```{r fig.width=15, fig.height=7}
forage_beta = ggarrange(pmain_bray_forage, pmain_wuF_forage, labels = c("A","B"), ncol = 2)
forage_beta
```

#Export beta-diversity plots of forage percentage as PDF
```{r}

setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Manning_lab/Microbiome_cattle/results/2021")
# mm to inch
setWidth = 15

# font size in pt
setFontSize = 7

# 1 in R = 0.75pt, so 0.25pt is specified as 
setLwd <- 0.25/0.75    


pdf(file='beta_div_forage.pdf',width=setWidth,height=4,pointsize=setFontSize)
forage_beta
dev.off()

```
